### 2023-02-01 manuscript.Rmd
add a lot

# Proposed Criteria
## ID 9: Citations
@gartner2022responsible suggest the number of citations (their indicator #9 in Table 3) to evaluate the impact of a software package.
However, citations suffer from serious shortcoming when used to evaluate the impact of research software.
For example, it is much more commonly accepted to cite data analysis packages than supporting packages (such as papaja [@R-papaja], RMarkdown, RStudio, and LaTeX used to render this article). Therefore, the total citation count is confounded with the type of software.
A further complication arises through the fact that functionality of successful, modular scientific software is imported, re-used, and re-exported in other software packages.
While we highly encourage this from both a software engineering perspective and for scientific progress, it poses further difficulties to citation-based metrics.
This especially punishes highly relied-on backends such as optimization algorithms: For example, the NLopt optimization suite is cited 1,711 times on google scholar.
However, the algorithms from that suite are used in countless software packages, for example, as a backend of OpenMx which again has 811 citations.
Moreover, the algorithms from that suite are all open source, and therefore serve as a reference for others.
As an example, consider the lme4 package, that adapted one of the algortihms from NLopt.
This package has 57,820 citations on google scholar, yet no single lme4 user probably ever cited the NLopt library.
We therefore argue that merrit statements are much more useful to assess impact of research software.

@@

## Open Source
is not used in phase 1

@@

We think that well-tested software implies systematic bug chasing and unit tests, and that the difference between "fairly extensive" and "extensive" documentation is unclear.
As a result, this indicator is more of a "gut-feeling" indicator, roughly assessing the "size" of the software project.
We instead propose to assess the different aspects of rigor separately, more sharply defined, and also assess "usage" indepentently from quality.

# Aspects of rigor for Software Implementations
Best practices of SE:
## unit testing
## CI
## documentation
## open source
## bug reporting and disclosure in changelogs
## activate maintenance
## reproducibility

>>>>

# Proposed Criteria

## ID 9: Citations
@gartner2022responsible suggest the number of citations (their indicator #9 in Table 3) to evaluate the impact of a software package.
However, citations suffer from serious shortcoming when used to evaluate the impact of research software.
For example, it is much more commonly accepted to cite data analysis packages than supporting packages (such as papaja [@R-papaja], RMarkdown, RStudio, and LaTeX used to render this article). Therefore, the total citation count is confounded with the type of software.
A further complication arises through the fact that functionality of successful, modular scientific software is imported, re-used, and re-exported in other software packages.
While we highly encourage this from both a software engineering perspective and for scientific progress, it poses further difficulties to citation-based metrics.
This especially punishes highly relied-on backends such as optimization algorithms: For example, the NLopt optimization suite is cited 1,711 times on google scholar.
However, the algorithms from that suite are used in countless software packages, for example, as a backend of OpenMx which again has 811 citations.
Moreover, the algorithms from that suite are all open source, and therefore serve as a reference for others.
As an example, consider the lme4 package, that adapted one of the algortihms from NLopt.
This package has 57,820 citations on google scholar, yet no single lme4 user probably ever cited the NLopt library.
We therefore argue that merrit statements are much more useful to assess impact of research software.

@@

## ID 7: Contributer Roles and Involvement
<!-- Larger software project, but the same effort of the individual researcher leads to lower scores? -->

## ID 8: License
Currently, whether the software is open source is not used in computing a treshold for phase one of the assessment procedure.
However, we think the license is central for the value of  software as a research product.
Above, we already discussed the different ways research software can have an impact - not only by direct usage, but also by reimporting and reusing software in other packages.
This type of impact is not possible for closed-source software.
In addition, many aspects of rigor are not possible to evaluate for closed-source software - for example, whether it is well-tested or whether bugs have been fixed.
We therefore propose to either score the license directly, giving points to open-source licenses as defined by the Open Source Society (https://opensource.org/licenses), or to assess the indicators of rigor as discussed in the second part of this commentary, where it is only possible to gain points for certain aspects of rigor if the software is open-source.

@@

We think that well-tested software implies systematic bug chasing and unit tests, and that the difference between "fairly extensive" and "extensive" documentation is unclear.
As a result, this indicator is more of a "gut-feeling" indicator, roughly assessing the "size" of the software project.
We instead propose to assess "usage" or impact indepentently from quality/rigor.
We also propose to assess the different aspects of rigor (testing, documentation, etc.) separately, and more sharply defined.
We therefore dicuss in the following aspects of rigor for research software, and how they could be assessed.

# Aspects of rigor for Software Implementations
Best-practises for software development can be used to assess rigor for research software.
In the following, we discuss aspects that should be scored in phase one of the assessment.
<!-- some introductory text missing -->
<!-- maybe cite
https://joss.readthedocs.io/en/latest/review_criteria.html
and
https://joss.readthedocs.io/en/latest/review_checklist.html
here -->
## Tests
Tests are essential to verify the correctness of scientific software.
They are usually separated into different categories:
*Unit tests* evaluate only small pieces of code, for example specific R-functions.
*Functional Tests* test whether the input of a user interacting with the software leads to the correct output.
There are many more types of tests, and the type of tests appropriate for a specific software product depends on the software itself.
However, we argue it should at least be tested whether the software can be installed on different platforms, and whether the core functionality of the software is executed without errors (*Smoke Test*).
In addition, it might be possible to give higher scores if different types of tests are employed, or to define a minimum of tests types that should be present, for example *unit tests*, *integration tests* and *functional tests*.
Another metric that could be assessed is code coverage, defined as the percentage of code lines that are executed during testing.
However, we would like to note that this metric is to be taken with a grain of salt, as the quality and number of the tests is not taken into account.

## CI
## Documentation

## open source
## bug reporting and disclosure in changelogs
## activate maintenance
## reproducibility
